---
title: "Sales Analytics"
author: "Fayez Alshehri"
date: "9/24/2020"
output:
  
  rmdformats::readthedown:
  html_document: 
    
    theme: rmdformats::material
  pdf_document: default
---
```{r setup, include=FALSE, cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
```{r, echo=FALSE,include=FALSE, cache=FALSE, results=FALSE, warning=FALSE, comment=FALSE, warning=FALSE}
library(gridExtra)
library(tidyverse)
library(wesanderson)
library(ggthemes)
library(plotly)
library(rmarkdown)
library("readxl")
library(tidyverse)
library(dplyr)
library(ggplot2)
library(plotly)

library(lubridate)


Apr <- read_excel('test.xlsx',sheet = 'Apr',col_names=TRUE)



Apr[1,1] <- 'Model'
colnames(Apr) <- Apr[1,]
Apr <- Apr[-1,]
long_apr <- t(Apr)
long_apr <- as.data.frame(long_apr)
colnames(long_apr) <- long_apr[1,]
long_apr <- long_apr[-1,]
long_apr <- sapply( long_apr, as.numeric )
long_apr <- as.data.frame(long_apr)

Apr_wide <- Apr
Apr <- long_apr

Apr <- data.frame(Date = row.names(Apr), Apr)
Apr <- gather(Apr, "Model", "Quantity", -Date)
Apr$Date <- as.numeric(Apr$Date)
### March Data

Mar <- read_excel('test.xlsx',sheet = 'Mar')

Mar <- Mar[,1:9]

colnames(Mar)<- c("Region","Channel","Name","AA","BB","CC","DD","EE","FF")
Mar <- Mar[-1,]

Mar$AA <- as.numeric(Mar$AA)
Mar$BB <- as.numeric(Mar$BB)
Mar$CC <- as.numeric(Mar$CC)
Mar$DD <- as.numeric(Mar$DD)
Mar$EE <- as.numeric(Mar$EE)
Mar$FF <- as.numeric(Mar$FF)

Mar[is.na(Mar)] <- 0 

### Feb Data 

Feb <- read_excel('test.xlsx',sheet = 'Feb')

Feb <- Feb[,1:9]

colnames(Feb)<- c("Region","Channel","Name","AA","BB","CC","DD","EE","FF")
Feb <- Feb[-1,]

Feb$AA <- as.numeric(Feb$AA)
Feb$BB <- as.numeric(Feb$BB)
Feb$CC <- as.numeric(Feb$CC)
Feb$DD <- as.numeric(Feb$DD)
Feb$EE <- as.numeric(Feb$EE)
Feb$FF <- as.numeric(Feb$FF)

Feb[is.na(Feb)] <- 0 

#### Jan

Jan <- read_excel('test.xlsx',sheet = 'Jan')

Jan <- Jan[,1:9]

colnames(Jan)<- c("Region","Channel","Name","AA","BB","CC","DD","EE","FF")
Jan <- Jan[-1,]

Jan$AA <- as.numeric(Jan$AA)
Jan$BB <- as.numeric(Jan$BB)
Jan$CC <- as.numeric(Jan$CC)
Jan$DD <- as.numeric(Jan$DD)
Jan$EE <- as.numeric(Jan$EE)
Jan$FF <- as.numeric(Jan$FF)

Jan[is.na(Jan)] <- 0 



############ Gather And Combine
Jan <- gather(Jan, "Model", "Quantity", -Region,-Channel,-Name)
Jan$Month <- "Jan"
Feb <- gather(Feb, "Model", "Quantity", -Region,-Channel,-Name)
Feb$Month <- "Feb"
Mar <- gather(Mar, "Model", "Quantity", -Region,-Channel,-Name)
Mar$Month <- "Mar"
df <- rbind(Jan,Feb,Mar)

df$Through <- 0
df$Price <- 0

df[df$Month=="Jan",]$Through <- 10
df[df$Month=="Feb",]$Through <- 30
df[df$Month=="Mar",]$Through <- 10

df[df$Model=="AA",]$Price <- 1
df[df$Model=="BB",]$Price <- 2
df[df$Model=="CC",]$Price <- 3
df[df$Model=="DD",]$Price <- 4
df[df$Model=="EE",]$Price <- 5
df[df$Model=="FF",]$Price <- 6

df <- df[,c(6,1,2,3,4,5,7,8)]

```

## Intro{.tabset .tabset-fade}

As part of Vivo Data Analytics screening, I was handed an simulated Excel Data and was challenged to provide Sales Insight, the Data consist of 2 main tables, first is a time series of April Sales, the second is sales summary of multiple products through different distributed channels. 

A look at the data after being processed for our analysis



### April TimeSeries
```{r, layout="l-body-outset",echo=FALSE,message=FALSE,warning=FALSE}
library(rmarkdown)
paged_table(Apr)
```

<br>
<br>
<br>
<br>
<br>
<br>

### First Quarter
```{r, layout="l-body-outset",echo=FALSE,message=FALSE,warning=FALSE}
paged_table(df)
```






<br>
<br>
<br>
<br>
<br>
<br>


## April Analysis

As we can see from the chart bellow, sales was higher at the beginning of April and during the last 2 days. FF Model accounted for 38% of Total Sales
```{r echo=FALSE}
AprTrends <- Apr %>% ggplot(aes(x=Date,y=Quantity,color=Model)) + geom_line()

ggplotly(AprTrends)
```

```{r ,echo=FALSE, message=FALSE,warning=FALSE}

dougnot <- Apr %>% group_by(Model) %>% summarise(Quantity=sum(Quantity)) %>%  mutate(Percentage = (Quantity/sum(Quantity) * 100))
  

dougnot$ymax = cumsum(dougnot$Percentage)

# Compute the bottom of each rectangle
dougnot$ymin = c(0, head(dougnot$ymax, n=-1))
dougnot$labelPosition <- (dougnot$ymax + dougnot$ymin) / 2
dougnot$label <- paste0(dougnot$Model, "\n Percentage: %", round(dougnot$Percentage,3))
# Make the plot
ggplot(dougnot, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Model)) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), size=6) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none")




```


<br>
<br>
<br>
<br>
<br>
<br>

## First Quarter


The Bar-Graph bellow shows the total quantity sold for each Model in the first Quarter, Febraury Had a huge spike in the sold quantities which positively correlates with Sales-Through of 30 per Model.

```{r,echo=FALSE,message=FALSE,warning=FALSE}





df$Month <- factor(df$Month,levels = c("Jan", "Feb", "Mar"))

monthly <- df %>% group_by(Month,Model) %>% summarise(Sum = sum(Quantity)) %>%
  ggplot(aes(x=Month,y=Sum,fill=Model)) + geom_col()

ggplotly(monthly)

```

